\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-01-29 3-advanced.Rmd
more work on matching in the advanced feature section

Temporal matching checks how spatially matched pairs align temporally. We use the following chart to illustrate how the temporal matching works:

For each spatially matched pair, say `A` and `a`, we first find the largest `n` points in each series, colored in brown points here. Here we use the largest three but you can tune this number by `temporal_n_highest`. Then we construct the interval of the largest points from one series and see how many points, from the other series, fall into the intervals. The series used to construct the interval is controlled by `temporal_independent` and the window size by `temporal_window` with a default of 5.

In this illustration, we construct the interval based on series `A` and two of the three peaks from `a` falls into this interval at Time 7 and 27.

There's another mandatory argument that hasn't been introduced above: `temporal_var_to_match`. This argument controls the variable to match and it needs to appear in both the `major` and `minor` set. In the water level matching example, we match the variable `Water_course_level` from `river` to `prcp` from `climate`, hence need to manually rename one of them to match the other, here we rename `Water_course_level` to `prcp` in `river`:

Figure \ref{fig:illu-matching}

```{r illu-matching, out.width="100%", out.height="30%", fig.cap="sdfasdf"}
knitr::include_graphics("figures/illu-matching.png")
```

>>>>

Data fusion is a process that consider combining multiple multiple data sources and it has been applied widely in the remote sensing community.

Meteorological and hydrological data have been matched in the applied studies in the literature to understand drought, ... [ref]. To be a valid pair of matching, the matched pair from different data sources need to

  - be spatially close to each other, and 
  - have similar temporal movement

`match_sites()` from `cubble` matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series `A` and `a`, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series `A` and the peaks in series `a` are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in `match_sites()` to fine-tune the matching:

<!-- [can also think about matching as a form of correlation between series from a statistics point of view.] -->

```{r illu-matching, out.width="100%", fig.cap="An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series A. Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches."}
knitr::include_graphics("figures/illu-matching.png")
```

 - `spatial_n_keep`: the number of spatial match for each site to keep
 - `spatial_dist_max`: the maximum distance allowed for a matched pair 
 - `temporal_n_highest`: the number of peak used - 3 in the example above
 - `temporal_window`: the length of the interval - 5 in the example above
 - `temporal_min_match`: the minimum number of matched peak for a valid matched pair

### 2022-01-29 main.tex
more work on matching in the advanced feature section

Temporal matching checks how spatially matched pairs align temporally.
We use the following chart to illustrate how the temporal matching
works:

For each spatially matched pair, say \texttt{A} and \texttt{a}, we first
find the largest \texttt{n} points in each series, colored in brown
points here. Here we use the largest three but you can tune this number
by \texttt{temporal\_n\_highest}. Then we construct the interval of the
largest points from one series and see how many points, from the other
series, fall into the intervals. The series used to construct the
interval is controlled by \texttt{temporal\_independent} and the window
size by \texttt{temporal\_window} with a default of 5.

In this illustration, we construct the interval based on series
\texttt{A} and two of the three peaks from \texttt{a} falls into this
interval at Time 7 and 27.

There's another mandatory argument that hasn't been introduced above:
\texttt{temporal\_var\_to\_match}. This argument controls the variable
to match and it needs to appear in both the \texttt{major} and
\texttt{minor} set. In the water level matching example, we match the
variable \texttt{Water\_course\_level} from \texttt{river} to
\texttt{prcp} from \texttt{climate}, hence need to manually rename one
of them to match the other, here we rename \texttt{Water\_course\_level}
to \texttt{prcp} in \texttt{river}:

Figure \ref{fig:illu-matching}

@@

{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/illu-matching}

@@

\caption[sdfasdf]{sdfasdf}\label{fig:illu-matching}
\end{figure}
\end{CodeChunk}

>>>>

Data fusion is a process that consider combining multiple multiple data
sources and it has been applied widely in the remote sensing community.

Meteorological and hydrological data have been matched in the applied
studies in the literature to understand drought, \ldots{} {[}ref{]}. To
be a valid pair of matching, the matched pair from different data
sources need to

\begin{itemize}
\tightlist
\item
  be spatially close to each other, and
\item
  have similar temporal movement
\end{itemize}

\texttt{match\_sites()} from \texttt{cubble} matches data based on these
two criteria. It first matches the two data sources spatially through
computing the pairwise distance based on latitude and longitude. Pairs
that pass the spatial matching are then matched temporally through
computing the number of matched peaks within a fixed length window.
Figure \ref{fig:illu-matching} illustrate this temporal matching in more
details. Given two series \texttt{A} and \texttt{a}, 3 peaks have been
picked in each series. An interval, with default length of 5, is
constructed for each peak in series \texttt{A} and the peaks in series
\texttt{a} are tested against whether they fall into the any of the
intervals. In this illustration, there are 2 matches for these two
series. Several arguments are available in \texttt{match\_sites()} to
fine-tune the matching:

@@

{\centering \includegraphics[width=1\linewidth]{figures/illu-matching}

@@

\caption[An illustration of temporal matching in cubble]{An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series A. Two peaks in series a fall into the intervals and hence the two series are considered to have 2 matches.}\label{fig:illu-matching}
\end{figure}
\end{CodeChunk}

\begin{itemize}
\tightlist
\item
  \texttt{spatial\_n\_keep}: the number of spatial match for each site
  to keep
\item
  \texttt{spatial\_dist\_max}: the maximum distance allowed for a
  matched pair
\item
  \texttt{temporal\_n\_highest}: the number of peak used - 3 in the
  example above
\item
  \texttt{temporal\_window}: the length of the interval - 5 in the
  example above
\item
  \texttt{temporal\_min\_match}: the minimum number of matched peak for
  a valid matched pair
\end{itemize}
